---
title: "Take-home Exercise 3: Predicting HDB Public Housing Resale Pricies using Geographically Weighted Methods"
date: "6 March 2023"
date-modified: "`r Sys.Date()`"
number-sections: true
format: html
execute: 
  echo: true
  eval: true
  warning: false
editor: visual
---

# Setting the Scene

Housing is an essential component of household wealth worldwide.
Buying a housing has always been a major investment for most people.
The price of housing is affected by many factors.
Some of them are global in nature such as the general economy of a country or inflation rate.
Others can be more specific to the properties themselves.
These factors can be further divided to structural and locational factors.
Structural factors are variables related to the property themselves such as the size, fitting, and tenure of the property.
Locational factors are variables related to the neighbourhood of the properties such as proximity to childcare centre, public transport service and shopping centre.

Conventional, housing resale prices predictive models were built by using [**Ordinary Least Square (OLS)**](https://en.wikipedia.org/wiki/Ordinary_least_squares) method.
However, this method failed to take into consideration that spatial autocorrelation and spatial heterogeneity exist in geographic data sets such as housing transactions.
With the existence of spatial autocorrelation, the OLS estimation of predictive housing resale pricing models could lead to biased, inconsistent, or inefficient results (Anselin 1998).
In view of this limitation, **Geographical Weighted Models** were introduced for calibrating predictive model for housing resale prices.

## The Task

In this take-home exercise, you are tasked to predict HDB resale prices at the sub-market level (i.e. HDB 3-room, HDB 4-room and HDB 5-room) for the month of January and February 2023 in Singapore.
The predictive models must be built by using by using conventional OLS method and GWR methods.
You are also required to compare the performance of the conventional OLS method versus the geographical weighted methods.

## Packages used:

```{r}
pacman::p_load(tidyverse, sf, sfdep, tmap)
```

## Data used:

-   Geospatial Data

-   List of structural variables

    1.  Area of the unit
    2.  Floor level
    3.  Remaining lease
    4.  Age of the unit
    5.  Main Upgrading Program (MUP) completed

-   List of locational factors

    1.  Proxomity to CBD
    2.  Proximity to eldercare
    3.  Proximity to foodcourt/hawker centres
    4.  Proximity to MRT
    5.  Proximity to park
    6.  Proximity to good primary school
    7.  Proximity to shopping mall
    8.  Proximity to supermarket
    9.  Numbers of kindergartens within 350m
    10. Numbers of childcare centres within 350m
    11. Numbers of bus stop within 350m
    12. Numbers of primary school within 1km

# Load Data into R

## Geospatial Data

```{r}
mpsz <- st_read(dsn = "data/geospatial",
                layer = "MP14_SUBZONE_WEB_PL")
```
