---
title: "In-class Exercise 11: Spatial Interaction Models (SIM)"
date: "27 March 2023"
date-modified: "`r Sys.Date()`"
number-sections: true
format: html
execute: 
  echo: true
  eval: true
  warning: false
editor: visual
---

# Getting Started

```{r}
pacman::p_load(tmap, sf, sp, caret, stplanr, reshape2, broom, tidyverse)
```

# Preparing the Flow Data

## Importing OD Data

Firstly, we will import the Passenger Volume by Origin Destination Bus stops data set downloaded from LTA Datamall by using read_csv() of readr package.

```{r, eval = FALSE}
odbus <- read_csv("data/aspatial/origin_destination_bus_202210.csv")
```

```{r, eval = FALSE}
odbus$ORIGIN_PT_CODE <- 
  as.factor(odbus$ORIGIN_PT_CODE)
odbus$DESTINATION_PT_CODE <- 
  as.factor(odbus$DESTINATION_PT_CODE)
```

```{r, eval = FALSE}
odbus7_9 <- odbus %>%
  filter(DAY_TYPE == "WEEKDAY") %>%
  filter(TIME_PER_HOUR >= 7 &
           TIME_PER_HOUR <= 9) %>%
  group_by(ORIGIN_PT_CODE,
           DESTINATION_PT_CODE) %>%
  summarise(TRIPS = sum(TOTAL_TRIPS))
```

```{r, eval = FALSE}
write_rds(odbus7_9, "data/rds/odbus7_9.rds")
```

```{r}
odbus7_9 <- read_rds("data/rds/odbus7_9.rds")
```

## Extracting the study area

```{r}
busstop <- st_read(dsn = "data/geospatial",
                   layer = "BusStop") %>%
  st_transform(crs = 3414)
```

```{r}
mpsz <- st_read(dsn = "data/geospatial",
                layer = "MPSZ-2019") %>%
  st_transform(crs = 3414)
```

```{r}
busstop_mpsz <- st_intersection(busstop, mpsz) %>%
  select(BUS_STOP_N, SUBZONE_C) %>%
  st_drop_geometry()
```

```{r}
od_data <- left_join(odbus7_9, busstop_mpsz,
                     by = c("ORIGIN_PT_CODE"="BUS_STOP_N")) %>%
  rename(ORIGIN_BS = ORIGIN_PT_CODE,
         ORIGIN_SZ = SUBZONE_C,
         DESTIN_BS = DESTINATION_PT_CODE)
```

Before continuing, it is a good practice to check for duplicate records

```{r}
duplicate <- od_data %>%
  group_by_all() %>%
  filter(n()>1) %>%
  ungroup()
```

```{r}
od_data <- unique(od_data)
```

## Combining Busstop and od_data

```{r}
od_data <- left_join(od_data, busstop_mpsz,
                     by = c("DESTIN_BS"= "BUS_STOP_N"))
```

```{r}
duplicate <- od_data %>%
  group_by_all() %>%
  filter(n()>1) %>%
  ungroup()
glimpse(duplicate)
```

```{r, eval = FALSE}
od_data <- unique(od_data)
```

```{r, eval = FALSE}
od_data <- od_data %>%
  rename(DESTIN_SZ = SUBZONE_C) %>%
  drop_na()
```

Save the output into an rds file format

```{r, eval = FALSE}
write_rds(od_data, "data/rds/od_data.rds")
```

```{r}
od_data <- read_rds("data/rds/od_data.rds")
```

# Visualising Geospatial

```{r}
tmap_mode("plot")
tmap_options(check.and.fix = TRUE)
  qtm(mpsz)
```

# Isolating SUBZONE_C into new DF

```{r}
mpsz<-mpsz[order(mpsz$SUBZONE_C),]
head(mpsz,10)
```

# Computing Distance matrix

```{r}
mpsz_sp <- as(mpsz, "Spatial")
```

```{r}
dist <- spDists(mpsz_sp)
dist
```

## Sorting the distance matrix to subzone

```{r}
sz_names <- mpsz$SUBZONE_C
```

```{r}
colnames(dist) <- paste0(sz_names)
rownames(dist) <- paste0(sz_names)
```

### Pivoting distance value by Subzone_C

```{r}
distPair <- melt(dist) %>%
  rename(dist = value)
head(distPair, 10)
```

### Updating intra-zonal distances

```{r}
distPair %>%
  filter(dist > 0) %>%
  summary()
```

```{r, eval = FALSE}
distPair$dist <- ifelse(distPair$dist ==0,
                        50, distPair$dist)
```

```{r, eval = FALSE}
distPair <- distPair %>% 
  rename(orig = Var1, 
         dest = Var2)
```

```{r, eval = FALSE}
write_rds(distPair, "data/rds/distPair.rds")
```

```{r}
flow_data <- od_data %>%
  group_by(ORIGIN_SZ, DESTIN_SZ) %>%
  summarize(TRIPS = sum(TRIPS))
```

```{r}
pop <- read_csv("data/aspatial/pop.csv")
```

```{r}
pop <- pop %>%
  left_join(mpsz,
            by = c("PA"="PLN_AREA_N",
                   "SZ" = "SUBZONE_N")) %>%
  select(1:6) %>%
  rename(SZ_NAME = SZ,
         SZ = SUBZONE_C)

```

```{r, eval = FALSE}
flow_data1 <- flow_data %>%
  left_join(distPair,
            by = c("ORIGIN_SZ"= "orig",
                   "DESTIN_SZ" = "dest"))
```

```{r, eval = FALSE}
# Origin version
flow_data1 <- flow_data1 %>%
  left_join(pop,
            by = c(ORIGIN_SZ = "SZ")) %>%
  rename(ORIGIN_AGE7_12 = AGE7_12,
         ORIGIN_AGE13_24 = AGE13_24,
         ORIGIN_AGE25_64 = AGE25_64) %>%
  select(-c(PA,SZ_NAME))
```

```{r, eval = FALSE}
flow_data1 <- flow_data1 %>%
  left_join(pop,
            by = c(DESTIN_SZ = "SZ")) %>%
  rename(DESTIN_AGE7_12 = AGE7_12,
         DESTIN_AGE13_24 = AGE13_24,
         DESTIN_AGE25_64 = AGE25_64) %>%
  select(-c(PA,SZ_NAME))
```

```{r, eval = FALSE}
write_rds(flow_data1, "data/rds/flow_data1.rds")
```

```{r}
flow_data1 <- read_rds("data/rds/flow_data1.rds")
glimpse(flow_data1)
```
